Towards Correct Legion

### 2017-08-11 correctnesshpc/paper.tex
Small changes to paper

\title{A Formal Specification of Legion}
\author{George Stelle \inst{1} \and Noah Evans \inst{2}}
\institute{Los Alamos National Laboratory \\ 
\email{stelleg@lanl.gov}
\and Sandia National Laboratories 
\email{nevans@sandia.gov}}

\begin{abstract}
Legion is a programming model that constrains effects in a way that ensures
serial semantics are preserved. We formalize the Legion semantics in a
mechanized proof assistant, Coq, and show how this enables one to reason
formally and prove theorems about the semantics. We also discuss how to extend
this work to prove that serial semantics are preserved and how given enough
time, one could extend the effort to prove an implementation is correct.
\end{abstract}

@@

That said, what good is a computation if we aren't relatively certain that the
answers it returns are correct? Presumably, these computations inform important
decisions about topics ranging from climate to nuclear safety. We should be as
confident as we can be that we aren't basing these decisions on bugs.

@@

Legion is a programming model for high performance computing that enables
reasoning about with the attractive property that it is
guaranteed to preserve serial semantics. It achieves this by dividing memory
into logical regions. This allows programmers to avoid the
burden

\subsection{Theorem Provers}

@@

@@

One common theme in formalizing systems is the discovery of holes in the
informal specification. For example, when formalizing the C language for the
CompCert project, Leroy et. al. found many cases where the specification was
incomplete. Note that this is seperate notion from undefined behavior. Undefined
behavior in a specification is well defined. A little bit like known unknowns
vs. unknown unknowns, gaps in a specification

In formalizing Legion, we discovered that while there is an impressive theory
ensuring that the side effects of function calls running concurrently preserve
serial semantics, there is no theory for what function calls can run
concurrently based on value dependencies. For example, consider the infamous
fibonacci sequence example:

@@

It is well known that the function calls require

>>>>

\title{Towards Correct Legion}
\author{George Stelle \inst{1} \and Noah Evans \inst{2}}
\institute{Los Alamos National Laboratory \\ 
\email{stelleg@lanl.gov}
\and Sandia National Laboratories 
\email{nevans@sandia.gov}}

\begin{abstract}
Legion is a programming model that uses logical regions of memory to reason about 
locality and concurrency. We formalize the Legion semantics in a mechanized
proof assistant, Coq, and show how this enables one to reason formally and
prove theorems about the semantics. We also discuss how to extend this work to
prove that serial semantics are preserved and how given enough time, one could
extend the effort to prove an implementation is correct.
\end{abstract}

@@

That said, what good is a fast computation if we aren't relatively certain that
the answers it returns are correct? Presumably, these computations inform
important decisions about topics ranging from climate to nuclear safety. We
should be as confident as possible that we aren't basing these decisions on
unexpected semantics.

@@

Legion is a programming model for high performance computing that enables
reasoning about locality and with the attractive property that it is guaranteed
to preserve serial semantics. It achieves this by dividing memory into logical
regions.  This allows programmers to avoid the
burden

@@

Coq is a dependently typed proof assistant. At its core is the Calculus of
Inductive Constructions, a small dependently typed programming language and
logic, capable of powerful abstractions for formal proofs. In the last 10 years 
great strides have been made in using it to both build systems and prove useful
properties about them \cite{compcert, certicoq, chlipala...}.

Specifications come in a wide range of formats. Many, like the OpenMP
specification, are very informal, primarily consisting of english. Others, like
Legion's previously, are written by hand in the using standard notation for
logical relations, etc. For this paper, when we refer to a \emph{formal}
specification, we are referring to a specification defined in a formal logic,
like Coq.

While semantics defined in papers are a huge improvement over natural language
specifications, there is still room for improvement. \cite{formalspec vs paper} has 
shown that even mathematical specifications and proofs in computer science
papers almost always have errors that would be caught by a formal proof. This is not to say,
of course, that formal specifications and proofs are infallible. On the contrary, it 
is quite easy to get a specification wrong. That said, often, if a there is a mistake in a spec
, the process of proving it formally will expose that mistake, allowing for its correction.

@@

One common theme in formalizing systems is the discovery of holes in the
corresponding informal specification. For example, when formalizing the C
language for the CompCert project, Leroy et. al. found many cases where the
specification was incomplete. Note that this is seperate notion from undefined
behavior. Undefined behavior in a specification is well defined. A little bit
like known unknowns vs. unknown unknowns, gaps in a specification

In formalizing Legion, we discovered that while there is an impressive theory
ensuring that the side effects of function calls running concurrently preserve
serial semantics, there is no theory for what function calls can run
concurrently based on value dependencies. For example, consider the infamous
Fibonacci example, which naively computes the \texttt{n}'th Fibonacci number:

@@

Of course
